Automated Verification of Stand-alone Solar Photovoltaic Systems: Optimal Sizing and Project Validation

### 2019-12-31 Tex/Old text/abstract.tex
Thesis update

\thispagestyle{plain}
\begin{center}
%    \Large
%    \textbf{Automated Verification Applied to Stand-alone Solar Photovoltaic Systems}
%    
%    \vspace{0.4cm}
%    \large
%    Optimal Sizing and Project Validation
%    
%    \vspace{0.4cm}
%    \textbf{Alessandro Trindade}
%    
    \vspace{0.9cm}
    \textbf{Abstract}
\end{center}
With decreasing costs and increasing performance, the deployment of renewable energy systems is now growing faster now than in the past decade. In 2017, for the first time, the number of people without access to electricity dipped below 1 billion, but trends in energy access still fall short of global goals. Particular attention is given to stand-alone solar photovoltaic systems in rural areas or where grid extension is unfeasible. Tools to evaluate or to size electrification projects are available, but they are based on simulations that do not cover all aspects of the design space. However, the use of formal methods to model and validate any system has grown with time, mainly to find bugs in sophisticated hardware and software systems: they aim to establish system correctness with mathematical rigor. The use of formal methods in electrical systems is a new subject, with published research spanning only in the last four years.
Moreover, the use of automated synthesis in order to obtain the optimal sizing of solar photovoltaic systems has never been done before. This thesis marks the achievement of two major goals: first, the application of software model checking to verify formally the design of a stand-alone solar photovoltaic system, including solar panel, charge controller, battery, inverter, and electric load; second, a sound and automated approach to obtaining optimal sizing of stand-alone photovoltaic systems using program synthesis. For the  formal verification, we used case studies from real photovoltaic systems deployed in five different sites, ranging from 700 W to 1,200 W, in order to evaluate the proposed approach and to compare it with a specialized simulation tool. Different verification tools are evaluated also, in order to compare performance and soundness. Data from practical applications show the effectiveness of our proposed approach, where specific conditions that lead to failures in a photovoltaic solar system are detailed only by the automated verification method.
In addition, for the use of program synthesis, we propose a variant of the counterexample guided inductive synthesis (CEGIS) approach. This variant has two phases linking the technical and the cost analysis. First, we synthesize a feasible candidate based on power reliability, but which may not attain the lowest cost. Second, the candidate is then verified iteratively with a lower bound cost via symbolic model checking. If the verification step succeeds, the lower bound is adjusted; if it fails, a counterexample provides the optimal solution. The proposed synthesis method is novel and unprecedented as it streamlines the design of photovoltaic systems. Experimental results using seven case studies demonstrate that our synthesis method can produce optimal system sizing within an acceptable run-time. We also present a comparison with a specialized simulation tool over real photovoltaic systems in order to show the effectiveness of our approach, which can provide a more detailed and accurate solution than the simulation tool.

\textit{Keywords}: Formal verification; automated verification; model checking; program synthesis; electrical systems; and solar photovoltaic systems.

>>>>

\thispagestyle{plain}
\begin{center}
%    \Large
%    \textbf{Automated Verification Applied to Stand-alone Solar Photovoltaic Systems}
%    
%    \vspace{0.4cm}
%    \large
%    Optimal Sizing and Project Validation
%    
%    \vspace{0.4cm}
%    \textbf{Alessandro Trindade}
%    
    \vspace{0.9cm}
    \textbf{Abstract}
\end{center}
With decreasing costs and increasing performance, the deployment of renewable energy systems is growing faster now than in the past decade. In 2017, for the first time, the number of people without access to electricity dipped below 1 billion, but trends in energy access still fall short of global goals. Particular attention is given to stand-alone solar photovoltaic systems in rural areas or where grid extension is unfeasible. Tools to evaluate or to size electrification projects are available, but they are based on simulations that do not cover all aspects of the design space. However, the use of formal methods to model and validate any system has grown with time, mainly to find bugs in sophisticated hardware and software systems: they aim to establish system correctness with mathematical rigor. The use of formal methods in electrical systems is a new subject, with published research spanning only the last four years. Moreover, the use of automated synthesis in order to obtain optimal sizing of solar photovoltaic systems has never been done before. This thesis marks the achievement of two major goals: first, the application of software model checking to verify formally the design of a stand-alone solar photovoltaic system, including solar panel, charge controller, battery, inverter, and electric load; second, a sound, automated approach to obtaining optimal sizing of stand-alone photovoltaic systems using program synthesis. For the formal verification, we used case studies from real photovoltaic systems deployed in five different sites, ranging from 700 W to 1.200 W, in order to evaluate the proposed approach and to compare  it with a specialized simulation tool. Different verification tools are evaluated also, in order to compare performance and soundness. Data from practical applications show the effectiveness of our proposed approach, where specific conditions that lead to failures in a photovoltaic solar system are detailed only by the automated verification method. In addition, for the use of program synthesis, we propose a variant of the counterexample guided inductive synthesis (CEGIS) approach. This variant has two phases linking the technical and the cost analysis. First, we synthesize a feasible candidate based on power reliability, but which may not attain the lowest cost. Second, the candidate is then verified iteratively with a lower bound cost via symbolic model checking. If the verification step succeeds, the lower bound is adjusted; if it fails, a counterexample provides the optimal solution. The proposed synthesis method is novel and unprecedented as it streamlines the design of photovoltaic systems. Experimental results using seven case studies demonstrate that our synthesis method can produce optimal system sizing within an acceptable run-time. We also present a comparison with a specialized simulation tool over real photovoltaic systems in order to show the effectiveness of our approach, which can provide a more detailed and accurate solution than the simulation tool.

\textit{Keywords}: Formal verification; automated verification; model checking; program synthesis; electrical systems; solar photovoltaic systems.

### 2019-12-31 Tex/Old text/chapters/chapteroptimization.tex
Thesis update

%---------------------------------------------------------------------------
\section{Optimization Case studies} 
%---------------------------------------------------------------------------

@@

These case studies were defined based on usual electrical load found in riverside communities of the Amazon State in Brazil~\cite{abs-1811-09438, Agrener2013}. With exception of case 7 that was idealized as a small town solution to support few lamps and an 12 kBTUs air-conditioner.

%---------------------------------------------------------------------------
\section{Objectives and Setup} 
%---------------------------------------------------------------------------

@@

Only case study $7$ led to a \textit{time out} result in CPAchecker, i.e., 
it was not solved within $240$ minutes. However, if it is removed 
this time out limitation from CPAchecker, the verifier is 
able to solve the optimization in $44.97$ hours. 
The violation (SAT result) indicated in Table~\ref{tab1} 
is the $assert$ of line $22$ from Algorithm~\ref{alg:verification-algorithm}. %The results were tested by manual PV sizing and were sound (\textit{RQ1}). %, linking a feasible technical solution with the lowest cost possible, considering the equipment that were inputted to the code.

@@

>>>>

%---------------------------------------------------------------------------
\section{Description of the case studies} 
%---------------------------------------------------------------------------

@@

These case studies were defined based on usual electrical load found in riverside communities of the Amazon State in Brazil~\cite{TrindadeCordeiro19, Agrener2013}. With exception of case 7 that was idealized as a small town solution to support few lamps and an 12 kBTUs air-conditioner.

%---------------------------------------------------------------------------
\section{Objectives and Setup}
\label{sec:synthesissetup} 
%---------------------------------------------------------------------------

@@

Only case study $7$ led to a \textit{time out} result in CPAchecker, i.e., 
it was not solved within $240$ minutes. This is depicted in Fig.~\ref{fig:cpaoptcase7}. However, if it is removed this time out limitation from CPAchecker, the verifier is 
able to solve the optimization in $44.97$ hours. 
The violation (SAT result) indicated in Table~\ref{tab1} 
is the $assert$ of line $22$ from Algorithm~\ref{alg:verification-algorithm}. %The results were tested by manual PV sizing and were sound (\textit{RQ1}). %, linking a feasible technical solution with the lowest cost possible, considering the equipment that were inputted to the code.

@@

\begin{figure}[h]
\includegraphics[width=1.0\textwidth]{CPAchecker_timeout_case7.png}
\centering
\caption{CPAchecker time out result for case 7 optimization (file CPALog.txt).}
\label{fig:cpaoptcase7}
\end{figure}

### 2019-12-31 Tex/Old text/chapters/chapterverification.tex
Thesis update

Moreover, the proposed flowchart of the automated verification method is illustrated in Fig.~\ref{fig:flowchartgeneral}. Those three steps are the high level description of how the automated verification, applied to solar PV systems, works.

@@

%We have performed five case studies to evaluate our proposed verification method: (a) four PV systems (three in series 325W PV panels, four 220 Ah batteries in a configuration with two series and two parallel with 48h autonomy, 700 W inverter with peak power of 1,600W, charge controller with MPPT with 35A/150V of capacity) deployed in four different houses in an indigenous community (GPS coordinates 2$^{o}$44'50.0"S 60$^{o}$25'47.8"W) situated nearby Manaus (Brazil), with each house having a different power demand (house 1 = 253 W, house 2 = 263 W, house 3 = 283 W, and house 4 = 501 W); and (b) one case concerning a system deployed as an individual system in Manaus (GPS coordinates 3$^{o}$4'20.208"S 60$^{o}$0'30.168"W), supporting 915 W of the house's load (house 5 with four 325W PV panels in a configuration two series and two parallel, four 120Ah batteries in series and autonomy of just 6 h, 1,200 W inverter with surge of 1,600 W, charge controller with MPPT of 150V/35A). 
It was performed five case studies to evaluate the proposed approach as described in Table~\ref{tab2}. %Furthermore, three start-of-art verification tools, as described in Section~\ref{sec:AutomatedVerification} (ESBMC, CBMC, and CPAchecker), and HOMER Pro simulation tool were used to compare the approach effectiveness and efficiency.
These case studies were defined based on usual electrical load found in riverside communities of the Amazon State in Brazil~\cite{abs-1811-09438, Agrener2013}.

@@

\subsection{ESBMC}

@@

\subsection{CPAchecker}

>>>>

Moreover, the proposed flowchart of the automated verification method is illustrated in Fig.~\ref{fig:flowchartgeneral}. Those three steps are the high level description of how the automated verification method can be used to validate solar PV systems.

@@

%We have performed five case studies to evaluate our proposed verification method: (a) four PV systems (three in series 325W PV panels, four 220 Ah batteries in a configuration with two series and two parallel with 48h autonomy, 700 W inverter with peak power of 1,600W, charge controller with MPPT with 35A/150V of capacity) deployed in four different houses in an indigenous community (GPS coordinates 2$^{o}$44'50.0"S 60$^{o}$25'47.8"W) situated nearby Manaus (Brazil), with each house having a different power demand (house 1 = 253 W, house 2 = 263 W, house 3 = 283 W, and house 4 = 501 W); and (b) one case concerning a system deployed as an individual system in Manaus (GPS coordinates 3$^{o}$4'20.208"S 60$^{o}$0'30.168"W), supporting 915 W of the house's load (house 5 with four 325W PV panels in a configuration two series and two parallel, four 120Ah batteries in series and autonomy of just 6 h, 1,200 W inverter with surge of 1,600 W, charge controller with MPPT of 150V/35A). 
It was performed five case studies to evaluate the proposed approach as described in Table~\ref{tab2}. %Furthermore, three start-of-art verification tools, as described in Section~\ref{sec:AutomatedVerification} (ESBMC, CBMC, and CPAchecker), and HOMER Pro simulation tool were used to compare the approach effectiveness and efficiency.
These case studies were defined based on usual electrical load found in riverside communities of the Amazon State in Brazil~\cite{TrindadeCordeiro19, Agrener2013}.

@@

\subsection{ESBMC}
\label{sec:ESBMCverification}

@@

\subsection{CPAchecker}
%\label{sec:CPAresult1}

### 2019-12-31 Tex/Old text/chapters/introduction.tex
Thesis update

The main objective of this dissertation is thus to create a new approach of formal methods and evaluate it as a solution to research issues pertaining to solar PV systems. We therefore adopt specific computer science techniques to tackle specific issues of PV systems, such as design validation and sizing optimization. In particular, we focus on automated verification and automated synthesis. Secondarily, we focus on: evaluating different state-of-the-art symbolic software verifiers (and solvers) with the goal of obtaining the best performance in our verification back-end;  applications written in ANSI-C that are platform-independent. We further evaluate our proposed solution comparing it with commercial simulation tools and using data collected from real PV systems deployed in riverside communities.
%The main objective of this thesis is thus to create a new approach of formal methods and evaluate it as an alternative to solve research issues from solar PV systems. We therefore foster the use of specific computer science techniques to tackle specific issues of PV systems, as design validation and sizing optimization.  In particular, we focus on automated verification and automated synthesis. Secondarily, we focus on: evaluating different state-of-art symbolic software verifiers (and solvers) with the goal of obtaining the best performance in our verification back-end; on applications written in ANSI-C that are platform-independent. %; and we do not perform code modification on the verifiers or solvers, we use it as available by their developers. 
%We further evaluate our proposed solution by comparative with commercial simulation tool and by data collected from real PV systems deployed at riverside communities.

>>>>

The main objective of this thesis is thus to create a new approach of formal methods and evaluate it as a solution to research issues pertaining to solar PV systems. We therefore adopt specific computer science techniques to tackle specific issues of PV systems, such as design validation and sizing optimization. In particular, we focus on automated verification and automated synthesis. Secondarily, we focus on: evaluating different state-of-the-art symbolic software verifiers (and solvers) with the goal of obtaining the best performance in our verification back-end;  applications written in ANSI-C that are platform-independent. We further evaluate our proposed solution comparing it with commercial simulation tools and using data collected from real PV systems deployed in riverside communities.
%The main objective of this thesis is thus to create a new approach of formal methods and evaluate it as an alternative to solve research issues from solar PV systems. We therefore foster the use of specific computer science techniques to tackle specific issues of PV systems, as design validation and sizing optimization.  In particular, we focus on automated verification and automated synthesis. Secondarily, we focus on: evaluating different state-of-art symbolic software verifiers (and solvers) with the goal of obtaining the best performance in our verification back-end; on applications written in ANSI-C that are platform-independent. %; and we do not perform code modification on the verifiers or solvers, we use it as available by their developers. 
%We further evaluate our proposed solution by comparative with commercial simulation tool and by data collected from real PV systems deployed at riverside communities.

### 2019-12-31 Tex/Old text/resumo.tex
Thesis update

Com custos decrescentes e com melhoria de desempenho, a implantação de sistemas de energia renovável está crescendo cada vez mais rapidamente no mundo. Em 2017, pela primeira vez, o número de pessoas sem acesso a eletricidade ficou abaixo de 1 bilhão, mas os dados quanto à universalização do acesso a energia ficaram aquém das metas globais. Particular atenção é dada aos sistemas isolados solares fotovoltaicos em áreas rurais ou onde a extensão da rede é inviável. 
Ferramentas para avaliar ou dimensionar projetos de eletrificação estão disponíveis, mas elas são baseadas em simulações que não cobrem todos os aspectos do espaço de projeto. Por outro lado, o uso de métodos formais para modelar e validar qualquer tipo de sistema está crescendo com o tempo, principalmente para encontrar bugs em sistemas complexos de \textit{hardware} e \textit{software}: seu objetivo é estabelecer a corretude do sistema com rigor matemático. 
O uso de métodos formais em sistemas elétricos é um assunto recente, com pesquisas sendo publicadas apenas nos últimos quatro anos. Além disso, a síntese automatizada nunca foi usada antes para obter um ótimo dimensionamento de sistemas solares fotovoltaicos nunca foi feito antes. 
Esta tese marca duas conquistas principais: (1) a primeira aplicação de verificação de modelos de software para verificar o projeto de um sistema isolado solar fotovoltaico, incluindo painel solar, controlador de carga, bateria, inversor e carga elétrica; e (2) uma abordagem confiável e automatizada para obter o dimensionamento ótimo de sistemas fotovoltaicos usando a síntese de programas onde cada componente e função de um sistema solar fotovoltaico é descrito, incluindo suas propriedades, e o modelo comportamental que representa o dimensionamento ótimo é sintetizado automaticamente. 
Relacionado à verificação formal, estudos de caso de sistemas fotovoltaicos reais instalados em cinco localidades diferentes são usados para avaliar a abordagem proposta e para compará-la com ferramenta de simulação especializada.
Diferentes ferramentas de verificação são avaliadas também, a fim de comparar o desempenho e a confiabilidade dos resultados. Dados de aplicações práticas mostram a eficácia da abordagem proposta, onde condições específicas que levam a falhas em um sistema solar fotovoltaico são detalhadas apenas pelo método de verificação automatizado. 
Além disso, em relação ao uso da síntese de programas, propõe-se uma variante do método de síntese indutiva guiada por contraexemplos (CEGIS), com duas fases bem definidas: primeiro, ele sintetiza o dimensionamento de sistemas fotovoltaicos baseados em confiabilidade de energia, mas que pode não alcançar o menor custo; segundo, a solução proposta é então verificada iterativamente com um limite inferior via verificação de modelo simbólico. Se a etapa de verificação não falhar, o limite inferior será ajustado; e se falhar, o contraexemplo é fornecido com o dimensionamento ótimo, vinculando assim a resposta técnica da primeira fase à análise de custo da segunda fase. Os dados de equipamentos comerciais de diferentes fabricantes são fornecidos ao mecanismo de síntese e as soluções candidatas são derivadas da análise financeira do dimensionamento obtido. O método de síntese proposto é novo e sem precedentes para simplificar o projeto de sistemas fotovoltaicos. Resultados experimentais usando sete estudos de caso mostram que o nosso método de síntese é capaz de produzir em um tempo de execução aceitável o dimensionamento ótimo do sistema fotovoltaico, e um comparativo com uma ferramenta de simulação especializada e sistemas fotovoltaicos reais mostra a eficácia da abordagem adotada.

>>>>

Com custos decrescentes e com melhoria de desempenho, a implantação de sistemas de energia renovável está crescendo cada vez mais rapidamente no mundo. Em 2017, pela primeira vez, o número de pessoas sem acesso a eletricidade ficou abaixo de 1 bilhão, mas os dados quanto à universalização do acesso a energia ficaram aquém das metas globais. Particular atenção é dada aos sistemas isolados solares fotovoltaicos em áreas rurais ou onde a extensão da rede é inviável. Ferramentas para avaliar ou dimensionar projetos de eletrificação estão disponíveis, mas elas são baseadas em simulações que não cobrem todos os aspectos do espaço de projeto. Por outro lado, o uso de métodos formais para modelar e validar qualquer tipo de sistema está crescendo com o tempo, principalmente para encontrar bugs em sistemas complexos de \textit{hardware} e \textit{software}: seu objetivo é estabelecer a corretude do sistema com rigor matemático. O uso de métodos formais em sistemas elétricos é um assunto recente, com pesquisas sendo publicadas apenas nos últimos quatro anos. Além disso, a síntese automatizada nunca foi usada antes para obter um ótimo dimensionamento de sistemas solares fotovoltaicos . Esta tese marca duas conquistas principais: (1) a primeira aplicação de verificação de modelos de \textit{software} para verificar o projeto de um sistema isolado solar fotovoltaico, incluindo painel solar, controlador de carga, bateria, inversor e carga elétrica; e (2) uma abordagem confiável e automatizada para obter o dimensionamento ótimo de sistemas fotovoltaicos usando a síntese de programas onde cada componente e função de um sistema solar fotovoltaico é descrito, incluindo suas propriedades, e o modelo comportamental que representa o dimensionamento ótimo é sintetizado automaticamente. Relacionado à verificação formal, estudos de caso de sistemas fotovoltaicos reais instalados em cinco localidades diferentes são usados para avaliar a abordagem proposta e para compará-la com ferramenta de simulação especializada. Diferentes ferramentas de verificação são avaliadas também, a fim de comparar o desempenho e a confiabilidade dos resultados. Dados de aplicações práticas mostram a eficácia da abordagem proposta, onde condições específicas que levam a falhas em um sistema solar fotovoltaico são detalhadas apenas pelo método de verificação automatizado. Além disso, em relação ao uso da síntese de programas, propõe-se uma variante do método de síntese indutiva guiada por contraexemplos (CEGIS), com duas fases bem definidas: primeiro, ele sintetiza o dimensionamento de sistemas fotovoltaicos baseados em confiabilidade de energia, mas que pode não alcançar o menor custo; segundo, a solução proposta é então verificada iterativamente com um limite inferior via verificação de modelo simbólico. Se a etapa de verificação não falhar, o limite inferior será ajustado; e se falhar, o contraexemplo é fornecido com o dimensionamento ótimo, vinculando assim a resposta técnica da primeira fase à análise de custo da segunda fase. Os dados de equipamentos comerciais de diferentes fabricantes são fornecidos ao mecanismo de síntese e as soluções candidatas são derivadas da análise financeira do dimensionamento obtido. O método de síntese proposto é novo e sem precedentes para simplificar o projeto de sistemas fotovoltaicos. Resultados experimentais usando sete estudos de caso mostram que o nosso método de síntese é capaz de produzir em um tempo de execução aceitável o dimensionamento ótimo do sistema fotovoltaico, e um comparativo com uma ferramenta de simulação especializada e sistemas fotovoltaicos reais mostra a eficácia da abordagem adotada.
%Com custos decrescentes e com melhoria de desempenho, a implantação de sistemas de energia renovável está crescendo cada vez mais rapidamente no mundo. Em 2017, pela primeira vez, o número de pessoas sem acesso a eletricidade ficou abaixo de 1 bilhão, mas os dados quanto à universalização do acesso a energia ficaram aquém das metas globais. Particular atenção é dada aos sistemas isolados solares fotovoltaicos em áreas rurais ou onde a extensão da rede é inviável. Ferramentas para avaliar ou dimensionar projetos de eletrificação estão disponíveis, mas elas são baseadas em simulações que não cobrem todos os aspectos do espaço de projeto. Por outro lado, o uso de métodos formais para modelar e validar qualquer tipo de sistema está crescendo com o tempo, principalmente para encontrar bugs em sistemas complexos de \textit{hardware} e \textit{software}: seu objetivo é estabelecer a corretude do sistema com rigor matemático. O uso de métodos formais em sistemas elétricos é um assunto recente, com pesquisas sendo publicadas apenas nos últimos quatro anos. Além disso, a síntese automatizada nunca foi usada antes para obter um ótimo dimensionamento de sistemas solares fotovoltaicos nunca foi feito antes. Esta tese marca duas conquistas principais: (1) a primeira aplicação de verificação de modelos de software para verificar o projeto de um sistema isolado solar fotovoltaico, incluindo painel solar, controlador de carga, bateria, inversor e carga elétrica; e (2) uma abordagem confiável e automatizada para obter o dimensionamento ótimo de sistemas fotovoltaicos usando a síntese de programas onde cada componente e função de um sistema solar fotovoltaico é descrito, incluindo suas propriedades, e o modelo comportamental que representa o dimensionamento ótimo é sintetizado automaticamente. Relacionado à verificação formal, estudos de caso de sistemas fotovoltaicos reais instalados em cinco localidades diferentes são usados para avaliar a abordagem proposta e para compará-la com ferramenta de simulação especializada. Diferentes ferramentas de verificação são avaliadas também, a fim de comparar o desempenho e a confiabilidade dos resultados. Dados de aplicações práticas mostram a eficácia da abordagem proposta, onde condições específicas que levam a falhas em um sistema solar fotovoltaico são detalhadas apenas pelo método de verificação automatizado. Além disso, em relação ao uso da síntese de programas, propõe-se uma variante do método de síntese indutiva guiada por contraexemplos (CEGIS), com duas fases bem definidas: primeiro, ele sintetiza o dimensionamento de sistemas fotovoltaicos baseados em confiabilidade de energia, mas que pode não alcançar o menor custo; segundo, a solução proposta é então verificada iterativamente com um limite inferior via verificação de modelo simbólico. Se a etapa de verificação não falhar, o limite inferior será ajustado; e se falhar, o contraexemplo é fornecido com o dimensionamento ótimo, vinculando assim a resposta técnica da primeira fase à análise de custo da segunda fase. Os dados de equipamentos comerciais de diferentes fabricantes são fornecidos ao mecanismo de síntese e as soluções candidatas são derivadas da análise financeira do dimensionamento obtido. O método de síntese proposto é novo e sem precedentes para simplificar o projeto de sistemas fotovoltaicos. Resultados experimentais usando sete estudos de caso mostram que o nosso método de síntese é capaz de produzir em um tempo de execução aceitável o dimensionamento ótimo do sistema fotovoltaico, e um comparativo com uma ferramenta de simulação especializada e sistemas fotovoltaicos reais mostra a eficácia da abordagem adotada.
